---
title: "Linear regression for Information and Comunication Technology (ICT) Skill"
author: Geiser C. Challco <geiser@alumni.usp.br>
comment: This file is automatically generate by Shiny-Statistic app (https://statistic.geiser.tech/)
         Author - Geiser C. Challco <geiser@alumni.usp.br>
         
         Shiny-Statistic is distributed in the hope that it will be useful,
         but WITHOUT ANY WARRANTY; without even the implied warranty of
         MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
         GNU General Public License for more details.
         
         You should have received a copy of the GNU General Public License.
         If not, see <https://www.gnu.org/licenses/>.
output:
  github_document:
    toc: true
  html_document:
    toc: true
  word_document:
    toc: true
fontsize: 10pt
---

```{r setup, include=FALSE}
## Install and Load Packages and functions
wants <- c('readxl','dplyr')
has <- wants %in% rownames(installed.packages())
if (any(!has)) install.packages(wants[!has])


get_data <- function(dat, grp, is.mean = F, extra.fields=c()) {
  library(dplyr)

  years <- sort(unique(dat$year))
  dat2 <- do.call(rbind, lapply(years, FUN = function(year) {
    idx <- which(dat$IncomeGroup == grp & dat$year == year & !is.na(dat$value))
    if (length(idx) == 0) return(NULL)
    values <- dat$value[idx]
    if (is.mean) {
      data.frame(year=as.integer(year), ICT.SKILL=mean(values))
    } else {
      toReturn <- data.frame(year=as.integer(year), ICT.SKILL=values)
      if (length(extra.fields)>0) {
        toReturn <- cbind(toReturn, dat[idx,extra.fields])
        names(toReturn) <- c("year","ICT.SKILL", extra.fields)
      }
      toReturn
    }
  }))

  return(dat2)
}


library(readxl)
library(dplyr)
library(aomisc)
```

```{r, include=FALSE, purl=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
defaultW <- getOption("warn")
options(warn = -1)
```

## Initial Variables and Data

```{r}
mtd.df <- read_excel("../data/ICT_and_Digital_Skill.xlsx", sheet = "metadata")
GCI.df <- read_excel("../data/ICT_and_Digital_Skill.xlsx", sheet = "Digital Skill value")
CIL.df <- read_excel("../data/ICT_and_Digital_Skill.xlsx", sheet = "ICILS (CIL)")
```

### getting GCI 4.0 (Digital Skills)

```{r}
gci.idx <-which(mtd.df$`Indicator Id`=="41400")
gci.val <- 100*(GCI.df$`2017` - mtd.df$min[gci.idx])/(mtd.df$max[gci.idx] - mtd.df$min[gci.idx])

dat <- data.frame(country=GCI.df$`Country Name`,value=gci.val,year=2017,IncomeGroup=GCI.df$IncomeGroup, indicator="GCI")


gci.val <- 100*(GCI.df$`2018` - mtd.df$min[gci.idx])/(mtd.df$max[gci.idx] - mtd.df$min[gci.idx])

dat <- rbind(
  dat,
  data.frame(country=GCI.df$`Country Name`,value=gci.val,year=2018,IncomeGroup=GCI.df$IncomeGroup, indicator="GCI")
)

gci.val <- 100*(GCI.df$`2019` - mtd.df$min[gci.idx])/(mtd.df$max[gci.idx] - mtd.df$min[gci.idx])

dat <- rbind(
  dat,
  data.frame(country=GCI.df$`Country Name`,value=gci.val,year=2019,IncomeGroup=GCI.df$IncomeGroup, indicator="GCI")
)
```

### getting CIL (ICILS)

```{r}
cli.idx <-which(mtd.df$`Indicator Id`=="CIL")
cli.val <- 100*(CIL.df$CIL - mtd.df$min[cli.idx])/(mtd.df$max[cli.idx] - mtd.df$min[cli.idx])

dat <- rbind(
  dat,
  data.frame(country=CIL.df$country,value=cli.val,year=CIL.df$year,IncomeGroup=CIL.df$IncomeGroup, indicator="CIL")
)

dat <- dat[!is.na(dat$IncomeGroup) & !is.na(dat$value),]


lmdls <- c()
smdls <- c()

income_grp <- unique(dat$IncomeGroup)
(income_grp <- income_grp[!is.na(income_grp)])
```

## Calculating data with mean values

```{r}
df <- get_data(dat, income_grp[1], is.mean = T)
for (i in 2:length(income_grp)) {
  grp = income_grp[i]
  df <- merge(df, get_data(dat, grp, is.mean = T), by="year", suffixes = c(income_grp[i-1], grp))
}
colnames(df) <- c("year",income_grp)

knitr::kable(df)
```

## Linear Regression

### Linear regression for high income countries

```{r}
grp = income_grp[2]
dat2 <- get_data(dat, grp, F, c("indicator"))

fit <- lm(ICT.SKILL ~ year, data=dat2)
smdls[[grp]] <- fit
summary(fit)
```

```{r, dpi=300, fig.width=12, fig.height=8}
pyear <- seq(min(dat2$year), 2030, 1)

inds <- list("CIL"=c(17,"red"), "GCI"=c(16,"blue"))
plot(c(), c(), xlim = c(min(pyear), max(pyear)), ylim = c(0,100),
     main = grp, xaxt='n',yaxt='n', xlab = "year", ylab = "ICT.SKILL")
axis(1, at = seq(min(pyear), max(pyear), 1), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)
axis(2, at = seq(0, 100, 10), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)
for (ind in names(inds)) {
  points(dat2$year[which(dat2$indicator==ind)], dat2$ICT.SKILL[which(dat2$indicator==ind)],
         pch=as.integer(inds[[ind]][1]), col=inds[[ind]][2], cex=0.5)
}
matlines(pyear, predict(fit, newdata=list(year=pyear), interval="confidence"), lwd=1.25)
legend("bottomright", legend=names(inds), col=c("red","blue","purple"), lty=0,  cex = 0.75, bg = "transparent", pch=c(17,16,18), box.lty=0)
```

### Linear regression for upper middle income countries

```{r}
grp = income_grp[1]
dat2 <- get_data(dat, grp, F, c("indicator"))

fit <- lm(ICT.SKILL ~ year, data=dat2)
smdls[[grp]] <- fit
summary(fit)
```

```{r, dpi=300, fig.width=12, fig.height=8}
pyear <- seq(min(dat2$year), 2030, 1)

inds <- list("CIL"=c(17,"red"), "GCI"=c(16,"blue"))
plot(c(), c(), xlim = c(min(pyear), max(pyear)), ylim = c(0,100),
     main = grp, xaxt='n',yaxt='n', xlab = "year", ylab = "ICT.SKILL")
axis(1, at = seq(min(pyear), max(pyear), 1), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)
axis(2, at = seq(0, 100, 10), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)
for (ind in names(inds)) {
  points(dat2$year[which(dat2$indicator==ind)], dat2$ICT.SKILL[which(dat2$indicator==ind)],
         pch=as.integer(inds[[ind]][1]), col=inds[[ind]][2], cex=0.5)
}
matlines(pyear, predict(fit, newdata=list(year=pyear), interval="confidence"), lwd=1.25)
legend("bottomright", legend=names(inds), col=c("red","blue","purple"), lty=0,  cex = 0.75, bg = "transparent", pch=c(17,16,18), box.lty=0)
```

### Linear regression for lower middle income countries

```{r}
grp = income_grp[4]
dat2 <- get_data(dat, grp, F, c("indicator"))

fit <- lm(ICT.SKILL ~ year, data=dat2)
smdls[[grp]] <- fit
summary(fit)
```

```{r, dpi=300, fig.width=12, fig.height=8}
pyear <- seq(min(dat2$year), 2030, 1)

inds <- list("CIL"=c(17,"red"), "GCI"=c(16,"blue"))
plot(c(), c(), xlim = c(min(pyear), max(pyear)), ylim = c(0,100),
     main = grp, xaxt='n',yaxt='n', xlab = "year", ylab = "ICT.SKILL")
axis(1, at = seq(min(pyear), max(pyear), 1), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)
axis(2, at = seq(0, 100, 10), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)
for (ind in names(inds)) {
  points(dat2$year[which(dat2$indicator==ind)], dat2$ICT.SKILL[which(dat2$indicator==ind)],
         pch=as.integer(inds[[ind]][1]), col=inds[[ind]][2], cex=0.5)
}
matlines(pyear, predict(fit, newdata=list(year=pyear), interval="confidence"), lwd=1.25)
legend("bottomright", legend=names(inds), col=c("red","blue","purple"), lty=0,  cex = 0.75, bg = "transparent", pch=c(17,16,18), box.lty=0)
```

### Linear regression for low income countries

```{r}
grp = income_grp[3]
dat2 <- get_data(dat, grp, F, c("indicator"))

fit <- lm(ICT.SKILL ~ year, data=dat2)
smdls[[grp]] <- fit
summary(fit)
```

```{r, dpi=300, fig.width=12, fig.height=8}
pyear <- seq(min(dat2$year), 2030, 1)

inds <- list("CIL"=c(17,"red"), "GCI"=c(16,"blue"))
plot(c(), c(), xlim = c(min(pyear), max(pyear)), ylim = c(0,100),
     main = grp, xaxt='n',yaxt='n', xlab = "year", ylab = "ICT.SKILL")
axis(1, at = seq(min(pyear), max(pyear), 1), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)
axis(2, at = seq(0, 100, 10), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)
for (ind in names(inds)) {
  points(dat2$year[which(dat2$indicator==ind)], dat2$ICT.SKILL[which(dat2$indicator==ind)],
         pch=as.integer(inds[[ind]][1]), col=inds[[ind]][2], cex=0.5)
}
matlines(pyear, predict(fit, newdata=list(year=pyear), interval="confidence"), lwd=1.25)
legend("bottomright", legend=names(inds), col=c("red","blue","purple"), lty=0,  cex = 0.75, bg = "transparent", pch=c(17,16,18), box.lty=0)
```

### Summary of linear regression with average values

```{r, dpi=300, fig.width=12, fig.height=8}
pyear <- seq(min(dat$year), 2030, 1)
plot(x=pyear, y=c(), xlim = c(min(pyear), max(pyear)), ylim=c(0,100), ylab = "ICT.SKILL", xlab = "year", xaxt='n',yaxt='n')
axis(1, at = seq(min(pyear),max(pyear), 1), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)
axis(2, at = seq(0, 100, 10), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)


colors <- c("green","blue","red","orange")
names(colors) <- income_grp
for (grp in income_grp) {
  gdat <- get_data(dat, grp, is.mean = T)
  points(x=gdat$year, y=gdat$ICT.SKILL, pch=16, col=colors[grp], xaxt='n', yaxt='n', cex=1)
  matlines(pyear, predict(smdls[[grp]], newdata=list(year=pyear), interval="confidence"), lwd=0.75)
}
legend("bottomright", legend=c("upper middle income","high income","low income","lower middle income"),
       col=colors, lty=0,  cex = 0.75, bg = "transparent", pch=16, box.lty=0)
```


```{r, dpi=300, fig.width=12, fig.height=8}
pyear <- seq(min(dat$year), 2030, 1)
plot(x=pyear, y=c(), xlim = c(min(pyear), max(pyear)), ylim=c(0,100), ylab = "ICT.SKILL", xlab = "year", xaxt='n',yaxt='n')
axis(1, at = seq(min(pyear),max(pyear), 1), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)
axis(2, at = seq(0, 100, 10), tck = 1, lty = 4, col = "lightgray", lwd = 0.75)

colors <- c("green","blue","red","orange")
names(colors) <- income_grp
for (grp in income_grp) {
  gdat <- get_data(dat, grp, is.mean = F)
  points(x=gdat$year, y=gdat$ICT.SKILL, pch=16, col=as.character(colors[grp]), xaxt='n', yaxt='n', cex=0.5)
  matlines(pyear, col = colors[grp], predict(smdls[[grp]], newdata=list(year=pyear), interval="confidence"), lwd=1.25)
}
legend("bottomright", legend=c("upper middle income","high income","low income","lower middle income"),
       col=colors, lty=0,  cex = 0.75, bg = "transparent", pch=16, box.lty=0)
```


```{r include=FALSE, echo=FALSE, purl=FALSE}
options(warn = defaultW)
#knitr::purl("code/regression_ict.Rmd", "code/regression_ict.R", documentation = 2)
```
